---
title: "Week 6 (Data Manipulation)"
author: Autumn Neil
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


mtcars <- data.frame(mtcars)
```{r}
mtcars <- mtcars %>% as_tibble
```

```{r}
mtcars %>% slice(1:5)
```


```{r}
mtcars %>% slice(28:32)
```

```{r}
mtcars %>% count()
```

```{r}
mtcars %>% select(mpg)
```

```{r}
mtcars %>% filter(cyl == 6) %>% select(mpg)
```

```{r}
mtcars %>% filter(cyl == 6)
```

```{r}
mtcars %>% filter(mpg > 25) %>% select(c("mpg", "cyl"))
```


```{r}
library(tidyverse)
diamonds <- data.frame(diamonds)
diamonds[1:5, ]
nrow(diamonds)
nrow(diamonds[diamonds$cut == "Very Good", ]) 
nrow(diamonds[diamonds$carat > "3.0", ]) 
diamonds[diamonds$color == "D", c("color","cut")]
mean(diamonds$price)
```

```{r}
library(tidyverse)
diamonds <- diamonds %>% data.frame
diamonds %>% slice(1:5)
diamonds %>% count()
diamonds %>% filter(cut == "Very Good") %>% count()
diamonds %>% filter(carat > 3.0) %>% count()
diamonds %>% filter(color == "D") %>% select(c("color", "cut"))
diamonds %>% summarize(mean(price))
```



```{r}
mtcars %>% group_by(cyl) %>% summarize(n())
```

```{r}
mtcars %>% group_by(cyl) %>% summarize(mean(mpg), mean(disp))
```

```{r}
diamonds %>% group_by(cut) %>% summarize(max(price), min(price))
```

```{r}
diamonds %>% group_by(color) %>% summarize(max(price), mean(price), min(price))
```